### 2022-09-03 latex/Chapters/Conclusion.tex
update

\chapter{Conclusion} % Main chapter title

@@

TBW

>>>>

\chapter{Conclusion \& Professional Issues} % Main chapter title

@@

\textbf{Learning}
Learnt about puzzles, and also even took the time to learn a couple of sequences of moves that are sufficient to solve \textbf{RC}, 2x2x2 in labout one minute, 3x3x3 in about 2 minutes.
\\
\\
\textbf{Implementation}
Implemented a bunch of solvers. Ideally should do in C++ for speed and multithreaded (esp for MCTS)
\\
\\
\textbf{DRL \& DQL}
Quite impressive, without too much effort, how well they perform. However, I feel to really get more out of them (higher dimensions in particular) would need careful tuning of the convergence criteria.
\\
\\
\textbf{Professional issues}
\\
\\
See \cite{ManyWorlds}
\\
\\
See \cite{https://doi.org/10.1111/1758-5899.12718} and \cite{Bostrom2014}

### 2022-09-03 latex/Chapters/Puzzles.tex
update

%-----------------------------------
%	SECTION 2
%-----------------------------------
\section{Rubiks' Cube}

\subsection{History}
The second puzzle I will focus on is the well known Rubiks' Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}). Since it was commericalised in 1980, it has kown a worldwide success: countless people are playing with the famous original 3x3x3 cube, as well as with countless variations, more or less difficult of it. Competitions nowadays bring together \textit{speedcubers} who have trained for years and memorized various algorithms to solve the Rubiks in astonishly effective and fast times (literally in seconds for the best ones). Interestingly, some of the principles used by speedcubers generalise to different dimensions. As an example, it is quite impressive to watch \textit{Cubastic} solve a 15x15x15 RC in two and a half hours (\cite{151515Rubiks}).

@@

\cite{RubiksChicago}
\\
\cite{RubiksRadu}
\\

>>>>

%-----------------------------------
%	SECTION 2
%-----------------------------------
\section{Rubik's Cube}

\subsection{History}
The second puzzle I will focus on is the well known Rubik's Cube (\textbf{RC}), invented in 1974 by Erno Rubik, originally known under the name of \textit{Magic Cube} (see \cite{RubiksWiki}). Since it was commericalised in 1980, it has kown a worldwide success: countless people are playing with the famous original 3x3x3 cube, as well as with countless variations, more or less difficult of it. Competitions nowadays bring together \textit{speedcubers} who have trained for years and memorized various algorithms to solve the Rubik's in astonishly effective and fast times (literally in seconds for the best ones). Interestingly, some of the principles used by speedcubers generalise to different dimensions. As an example, it is quite impressive to watch \textit{Cubastic} solve a 15x15x15 RC in two and a half hours (\cite{151515Rubiks}).

@@

\cite{RubiksGodNumber}
\\
\cite{RubiksChicago}
\\
\cite{RubiksRadu}
\\

%-----------------------------------
%	SECTION 3
%-----------------------------------
\section{Solving puzzles \textit{without human knowledge}}

From handcrafted (naive, kociemba) to supervised (DL) to unsupervised (DRL \& DQL) .... but sometimes subtle knowledge injected in e.g. representation or simplifications.

### 2022-09-03 latex/Chapters/SlidingPuzzle.tex
update

 I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension). Running these experiments takes a lot of time, so I had to be somehow selective. I at least wanted to run all of the solvers I have implemented on the intermediary 3x3 dimension, and answer some of the following interesting questions:
\begin{itemize}
\item How hard is too hard for \textbf{BFS} to complete in \textit{finite} time?
\item How much, if any, improvement do we get with Manhattan++ over Manhattan?
\item Is there much loss between \textbf{DL} and \textbf{DRL}, i.e. is losing the teacher used by \textbf{DL} a deal breaker?
\item Is \textbf{DQL}, all things being equal, able to learn a better cost-to-go heuristic than \textbf{DRL}?
\item How does \textbf{MCTS} perform? How important is it to tune the hyper-parameter $c$? How much improvement, if any, does the post trim of the \textbf{MCTS} tree via \textbf{BFS} bring?
\item Are any of the \textbf{D*L} techniques able to perform on par with A$^{*}$ with Manhattan++ or perfect heuristic?
\end{itemize}
All of these questions are answered (obviously not in generality, but in the context of the experiments I have been running) in this chapter.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.63]{./Figures/44SPPerformanceAll}
%\decoRule
\caption[44SPPerformance]{Solvers' performance comparison 4x4 \textbf{SP}}
\label{fig:44SPPerformance}
\end{figure}

>>>>

I am neither making claims of depth (each solver could surely be tuned and optimized/adapted, possibly differently for each dimension) nor breadth (I have not run all of the solvers versus each dimension). Running these experiments takes a lot of time, so I had to be somewhat selective. I wanted however to run all of the solvers I have implemented one particular dimension (the intermediary 3x3 seemed appropriate to do so), to answer some of the following interesting questions:
\begin{itemize}
\item How hard is too hard for \textbf{BFS} to complete in \textit{reasonable} time?
\item How much, if any, improvement do we get with Manhattan++ over Manhattan?
\item Is there much loss between \textbf{DL} and \textbf{DRL}, i.e. is losing the teacher used by \textbf{DL} a deal breaker?
\item Is \textbf{DQL}, all things being equal, able to learn a better cost-to-go heuristic than \textbf{DRL}?
\item How does \textbf{MCTS} perform? How important is it to tune the hyper-parameter $c$? How much improvement, if any, does the post trim of the \textbf{MCTS} tree via \textbf{BFS} bring?
\item Are any of the \textbf{D*L} techniques able to perform on par with A$^{*}$ with Manhattan++ or perfect heuristic?
\end{itemize}
All of these questions are answered (obviously not in generality, but in the context of the experiments I have been running) in this chapter.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.63]{./Figures/44SPPerformance}
%\decoRule
\caption[44SPPerformance]{Solvers' performance comparison 4x4 \textbf{SP}}
\label{fig:44SPPerformance}
\end{figure}

### 2022-09-03 latex/MScAIFrancoisBerrier.tex
update

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters
\usepackage{mathpazo} % Use the Palatino font by default
\usepackage[T5,T1]{fontenc}
\usepackage{verbatim}
\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
\usepackage{float}
\usepackage{graphbox}
\usepackage{lscape}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[paper=portrait,pagesize]{typearea}
\usepackage{logicpuzzle}
\usepackage{geometry}
\newenvironment{three}[1][]{%
\begin{logicpuzzle}[rows=2,columns=2,#1]
\begin{puzzleforeground}
\framepuzzle
\end{puzzleforeground}
}{\end{logicpuzzle}}

@@

\thesistitle{Solving \\ the \\ 15-Puzzle \& Rubik's Cube \\by \\Deep Reinforcement Learning} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{\textbf{MSc in Artificial Intelligence}} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Fran\c{c}ois \textsc{Berrier}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

@@

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
\textbf{Motivation}
\\
Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in the Sutton book (\cite{Sutton1998}), has until recently known less success than we might have hoped for. Its framework is very appealing and intuitive. In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound. Sadly, it had until recently proven hard to achieve practical success with these methods.
\\
Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
\\
In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using (\textbf{DL}) or (\textbf{DRL}), to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
\\
\\
\textbf{Organisation of this thesis}
\\
In the first chapter, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!). In chapter 2, I will do a quick recap of the different methods that I will use in the project. Chapter 3 will be dedicated to discussing the mathematics of the sliding puzzle (\textbf{SP}) and the (\textbf{RC}). I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader. I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project before detailing a few examples in chapter 5. Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in chapter 8.

@@

\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
\begin{comment}
I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway. I entered in the study of AI with a heavy dose of skepticism and am delighted to have learnt so much and revised my judgment of this field.
In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for supporting me in pursuing this MSc.
\\
\\
I have been fortunate to have had the company of my colleague Saurabh Kumar, as he decided to follow me in this MSc journey. We have had countless opportunities to debate about all the new cool stuff we learnt during the MSc, and that has definitely made the whole process much more enjoyable.
\\
\\
I would also like to thank Professor Chris Watkins, inventor of QL :), for agreeing to supervise my project. It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our meetings during my work on this project.
\\
\\
Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 60 miles a week in preparation of multiple marathons and half marathons, and the coursework, lectures and exam revisions.
\end{comment}
\end{acknowledgements}
%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

@@

\textbf{AIPnT} & \textbf{A}rtificial \textbf{I}ntelligence \textbf{P}rinciples and \textbf{T}echniques\\
\textbf{BFS} & \textbf{B}readth \textbf{F}irst \textbf{S}earch\\
\textbf{CS} & \textbf{C}omputer \textbf{S}cience\\
\textbf{CV} & \textbf{C}omputer \textbf{V}ision\\
\textbf{DFS} & \textbf{D}epth \textbf{F}irst \textbf{S}earch\\
\textbf{DL} & \textbf{D}eep \textbf{L}earning\\
\textbf{DQL} & \textbf{D}eep \textbf{Q}-\textbf{L}earning\\
\textbf{DRL} & \textbf{D}eep \textbf{R}einforcment \textbf{L}earning\\
\textbf{ML} & \textbf{M}achine \textbf{L}earning\\
\textbf{NLP} & \textbf{N}atural \textbf{L}anguage \textbf{P}rocessing\\
\textbf{RC} & \textbf{R}ubiks' \textbf{C}ube\\
\textbf{RL} & \textbf{R}einforcement \textbf{L}earning\\
\textbf{RHUL} & \textbf{R}oyal \textbf{H}olloway, \textbf{U}niversity of \textbf{L}ondon\\
\textbf{SP} & \textbf{S}liding \textbf{P}uzzle\\

@@

@@

\include{Chapters/Objectives}
\include{Chapters/DRLSearch}
\include{Chapters/Puzzles}
\include{Chapters/Implementation}
\include{Chapters/Examples}
\include{Chapters/SlidingPuzzle}
\include{Chapters/Rubiks}
\include{Chapters/Conclusion}

@@

%\include{Appendices/AppendixA}
%\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}

@@

%\begin{thebibliography}{widest entry}

%\bibitem{sliddingpuzzle} Whatever
%{\href{https://en.wikipedia.org/wiki/Sliding\_puzzle}{wikipedia}} Sliding Puzzle Wikipedia article, accessed 19th June 2022

%\end{thebibliography}

>>>>

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters
\usepackage{mathpazo} % Use the Palatino font by default
\usepackage[T5,T1]{fontenc}
\usepackage{verbatim}
\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
\usepackage{float}
\usepackage{graphbox}
\usepackage{lscape}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[paper=portrait,pagesize]{typearea}
\usepackage{logicpuzzle}
\usepackage{geometry}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{rubikcube}
\newenvironment{three}[1][]{%
\begin{logicpuzzle}[rows=2,columns=2,#1]
\begin{puzzleforeground}
\framepuzzle
\end{puzzleforeground}
}{\end{logicpuzzle}}

@@

\thesistitle{Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Pr. Chris \textsc{Watkins}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{\textbf{MSc in Artificial Intelligence}} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Fran\c{c}ois \textsc{Berrier}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

@@

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
\textbf{Motivation}
\label{Abstract}
\\
Reinforcement Learning (\textbf{RL}), exposed with brilliant clarity in the Sutton book (\cite{Sutton1998}), has until recently known less success than we might have hoped for. Its framework is very appealing and intuitive. In particular, the mathematical beauty of Value iteration and Q iteration (\cite{WatkinsThesis}) for discrete state and action spaces, blindly iterating from \textit{any} initial value, is quite profound. Sadly, it had until recently proven hard to achieve practical success with these methods.
\\
Inspired however by the seminal success of Deep Mind's team in using Deep Reinforcement Learning (\textbf{DRL}) to play Atari games (\cite{Mnih2013})  and to master the game of Go (\cite{AlphaGo}), researchers have in recent years made a lot of progress towards designing algorithms capable of learning and solving, \textit{without human knowledge}, the Rubik's Cube (\textbf{RC}) - as well as similar single player puzzles - by using Deep Q-Learning (\textbf{DQL})  (\cite{DBLP:journals/corr/abs-1805-07470}) or search and (\textbf{DRL}) value iteration (\cite{https://doi.org/10.48550/arxiv.1805.07470}).
\\
In this project, I will attempt to implement a variety of solvers combining $A^{*}$ search and heuristics, some of which will be handcrafted, others which I will train on randomly generated sequences of puzzles using (\textbf{DL}) or (\textbf{DRL}), to solve the 15-puzzle (and variations of different dimensions), as well as the Rubik's cube.
\\
\\
\textbf{Organisation of this thesis}
\\
In the first chapter, I will quickly describe what I hope to get out of this project, both in terms of personal learning, as well as in terms of tangible results (solving some puzzles!). In chapter 2, I will do a quick recap of the different methods that I will use in the project. Chapter 3 will be dedicated to discussing the mathematics of the sliding puzzle (\textbf{SP}) and the (\textbf{RC}). I might throw a few random (but hopefully interesting) observations in there and give some references for the keen reader. I will then give an overview in chapter 4 of the code base I have developed - and put in the open on my github page (\cite{FB}) - to complete this project before detailing a few examples in chapter 5. Finally, in chapters 6 and 7, I will present all my various results on respectively the sliding puzzle and the Rubik's cube before offering some conclusion in chapter 8.

@@

\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
%\begin{comment}
I am grateful to my employer Bank of America for sponsoring my part time 2-year MSc in Artificial Intelligence at Royal Holloway. I entered in the study of AI with a heavy dose of skepticism and am delighted to have learnt so much and revised my judgment of this field.
In particular I would like to thank my managers Mitrajit Dutta and Stephen Thompson for supporting me in pursuing this MSc.
\\
\\
I have been fortunate to have had the company of my colleague Saurabh Kumar, as he decided to follow me in this MSc journey. We have had countless opportunities to debate about all the new cool stuff we learnt during the MSc, and that has definitely made the whole process much more enjoyable.
\\
\\
I would also like to thank Professor Chris Watkins, inventor of QL :), for agreeing to supervise my project. It was an honour to attend his very lively lectures on Deep Learning during the first year of the MSc, as well as to absorb some of his wisdom and enthusiasm at our meetings during my work on this project.
\\
\\
Last but not least, my wife Thanh Nh\~a is my biggest supporter in everything I do, and her continued encouragement and enthusiasm in my pursuit of this MSc has made it so much easier to juggle between a demanding banking job, running 50 to 70 miles a week in preparation for the Chicago marathon, and the coursework, lectures and exam revisions.
%\end{comment}
\end{acknowledgements}
%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

@@

\textbf{AIPnT} & \textbf{A}rtificial \textbf{I}ntelligence \textbf{P}rinciples and \textbf{T}echniques\\
\textbf{BFS} & \textbf{B}readth \textbf{F}irst \textbf{S}earch\\
\textbf{CS} & \textbf{C}omputer \textbf{S}cience\\
\textbf{CV} & \textbf{C}omputer \textbf{V}ision\\
\textbf{DFS} & \textbf{D}epth \textbf{F}irst \textbf{S}earch\\
\textbf{DL} & \textbf{D}eep \textbf{L}earning\\
\textbf{DQL} & \textbf{D}eep \textbf{Q}-\textbf{L}earning\\
\textbf{DRL} & \textbf{D}eep \textbf{R}einforcment \textbf{L}earning\\
\textbf{FIFO} & \textbf{F}irst \textbf{I}n \textbf{F}irst \textbf{O}ut\\
\textbf{LIFO} & \textbf{L}ast \textbf{I}n \textbf{F}irst \textbf{O}ut\\
\textbf{MCTS} & \textbf{M}onte \textbf{C}arlo \textbf{T}ree \textbf{S}earch \\
\textbf{ML} & \textbf{M}achine \textbf{L}earning\\
\textbf{NLP} & \textbf{N}atural \textbf{L}anguage \textbf{P}rocessing\\
\textbf{RC} & \textbf{R}ubik's \textbf{C}ube\\
\textbf{RL} & \textbf{R}einforcement \textbf{L}earning\\
\textbf{RHUL} & \textbf{R}oyal \textbf{H}olloway, \textbf{U}niversity of \textbf{L}ondon\\
\textbf{SP} & \textbf{S}liding \textbf{P}uzzle\\

@@

\hypersetup{%
  colorlinks = true,
  linkcolor  = orange
}

@@

\include{Chapters/Objectives}
\include{Chapters/DRLSearch}
\include{Chapters/Puzzles}
\include{Chapters/Implementation}
\include{Chapters/SlidingPuzzle}
\include{Chapters/Rubiks}
\include{Chapters/Conclusion}

@@

\include{Appendices/Examples}
\include{Appendices/Kociemba}

@@
